Using Fundamental Measure Theory to Treat the Correlation Function of the Inhomogeneous Hard-Sphere Fluid

### 2012-08-15 papers/contact/paper.tex
made some minor changes to appendices of paper - brought down twos from derives Tested everything on node06.

When we consider an inhomogeneous system, we may prefer to ask for
\emph{which} sphere we are interested in computing the contact
density.  One way to answer this is to specify a sphere at a fixed
position $\mathbf{r}$.  By making the hard-sphere radius dependent on
position, we can compute the pressure on the hard spheres located at
$\mathbf{r}$.  Since the hard-sphere radius enters the FMT free energy
by means of the fundamental measures, this requires that we adjust the
convolutions to have a radius $R$ which depends on the location of the
sphere:
\begin{align}
  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr')/2)d\rr' \label{eq:n3A}
\end{align}
Once we have done this---and similarly for the other fundamental
measures---we derive a formula for the correlation function
$g_\sigma^A(\rr)$ by finding the pressure on the surface of any hard
spheres located at position $\rr$, and using the contact-density
theorem to extract the contact density, from which we can establish
the averaged correlation function at contact.
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
The functional derivative $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
in the above formula must be computed with the asymmetric formulation
for the fundamental measures.  In practice, we use the approximate
White Bear functional to find an approximate expression for
$g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
expression will involve convolutions of derivatives of the free
energy, making this formulation computationally more expensive than
the free energy itself.

In addition to the different densities in the denominators of Equations 
\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
between the treatment of the A and S cases in the derivatives 
$2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
that we use involve a number of weighted densities and for the derivatives 
of these we use two different conceptual approaches.  In the assymetric case, 
a change in $\sigma(\mathbf{r})/2$ will refer to a change in the radius of a sphere 
that is located at the position $\mathbf{r}$.  The weighted densities that are 
affected by such a change will be those that are located a distance $\sigma/2$ away 
from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.

@@

These qualitative concepts lead us to the derivative terms in $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the
derivation of the correlatation function results when we recognize that
instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$
as in the assymmetric case we will divide by $n_0(\mathbf{r})^2$
since we have symmetry around the point $\mathbf{r}$.

\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^S(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
where the functional derivative $2 \frac{\delta A_{HS}}{\delta
  \sigma(\mathbf{r})}$ in this case is computed with the symmetric
formulation of the fundamental measures.  Complete expressions for
$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\

@@

In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.  We break these terms up as follows:

@@

We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.  We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.  Following this we will compute the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.  These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^A(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position of the sphere in
question---which is to say, the location where the density itself is
evaluated.  This is the most natural way to construct such a
functional, in which the radius of spheres is a function of position.
Thus we define $n_3^A$ as
\begin{align}
  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr'')/2) d\rr''
\end{align}
and we define the other weighted densities similarly.  In order to
compute derivatives of the free energy with respect to $R(\rr)$, we
need to find $g_\sigma^A$.  In the case of $n_3^A$, we needS
\begin{align}
  2 \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
  \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
   &= n (\rr) \delta(|\rr' - \rr| - \sigma(\rr)/2)
\end{align}
Similarly, for $n_2^A$, we find
\begin{align}
  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) d \rr''\\
  2 \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
\end{align}
This is somewhat more challenging, since it involves an unfamiliar
function, the derivative the Dirac delta function $\delta'$.
%
The derivatives of the remaining scalar densities can be reduced to
sums of the terms above:
\begin{align}
  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
  2 \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
  &= -\frac{n(\rr)}{2\pi
    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
  -
  \frac{n(\rr)}{\pi
    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2)
\end{align}
and
\begin{align}
  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
  2 \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
  &= -\frac{n(\rr)}{\pi
    \sigma(\rr)^2}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
  -
  4 \frac{n(\rr)}{\pi
    \sigma(\rr)^3}\delta(|\rr'-\rr| - \sigma(\rr)/2)
\end{align}
The vector-weighted densities have no additional occurrences of $R$,
their functional derivatives come out similarly:
\begin{align}
  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(|\rr' - \rr''| - \sigma(\rr'')/2)
    \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
  2 \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
    \frac{\rr-\rr'}{|\rr-\rr'|}\\
\end{align}
and analogously for $\mathbf{n}_{V1}^A$:
\begin{align}
  \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
  \delta(|\mathbf{r'}-\rr''| - \sigma(\rr'')/2) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
  2 \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
  &= \frac{n(\rr)}{2\pi
    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
  +
  \frac{n(\rr)}{\pi
    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
\end{align}
\fixme{double-check the directions of all the vector quantities...}

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^S(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position at which the weighted
density is being evaluated.  This is less intuitive than the $A$ case,
since \emph{nothing is there}!  However, in practice it corresponds to
the picture shown in Figure~\ref{fig:contact}, which involves
extending the radius for spheres just at the position touching a
given spot, where we are evaluating the correlation function.  We
define the weighted density $n_3^S$ as:
\begin{align}
  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr')/2) d\rr''
\end{align}
In the symmetric case, the functional derivative is strikingly simple:
\begin{align}
  2 \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
   \delta(\rr'-\rr) d\rr''
   \\
   &= \delta(\rr'-\rr) n_2^S(\rr')
\end{align}
The functional derivative does \emph{not} eliminate the integral,
which simply changes from one weighted density to another.  The
corresponding term in the free energy functional, once integrated,
takes a simple form:
\begin{align}
  \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
  \frac{\delta n_3^{S}(\rr')}{\delta \sigma(\rr)} d\rr'
  &= \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
  n_2^S(\rr') d\rr' \\
  &= 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
\end{align}
which means that we can compute $g_\sigma^S(\rr)$ using only
convolutions of the density, in strong contrast with
$g_\sigma^A(\rr)$, which requires convolutions of derivatives of the
free energy density.

The functional derivative of $n_2^S(\rr')$ is quite similar:
\begin{align}
  n_2^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf
  r''\\
  2 \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
  \delta'(|\rr'-\rr''| - \sigma(\rr')) d\rr''
\end{align}
where we recognize that the convolution---while not one of the standard
FMT weighted density---is a simple convolution of the density.  We can
go on to compute the derivatives of the remaining weighted densities:
\begin{align}
  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
    2\frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{2\pi \sigma(\rr)}
    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
    + n_0^s(\rr') \right)
\end{align}
\begin{align}
  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
    2 \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
    +
    \frac{4}{\sigma} n_0^S(\rr') \right)
\end{align}
\begin{align}
  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
  \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
  2 \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
  \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
  \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
\end{align}
\begin{align}
 \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''
   \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
 2 \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
 &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{2\pi \sigma(\rr)}
   \delta'(|\rr'-\rr''| - \sigma(\rr')/2) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
   + \frac{\mathbf{n}_{V2}^S(\rr')}{\pi \sigma(\rr)^2} \right)
\end{align}

>>>>

When we consider an inhomogeneous system, we may prefer to ask for
\emph{which} sphere we are interested in computing the contact
density.  One way to answer this is to specify a sphere at a fixed
position $\mathbf{r}$.  By making the hard-sphere radius dependent on
position, we can compute the pressure on the hard spheres located at
$\mathbf{r}$.  Since the hard-sphere radius enters the FMT free energy
by means of the fundamental measures, this requires that we adjust the
convolutions to have a radius $R$ which depends on the location of the
sphere:
\begin{align}
  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr')/2)d\rr' \label{eq:n3A}
\end{align}
Once we have done this---and similarly for the other fundamental
measures---we derive a formula for the correlation function
$g_\sigma^A(\rr)$ by finding the pressure on the surface of any hard
spheres located at position $\rr$, and using the contact-density
theorem to extract the contact density, from which we can establish
the averaged correlation function at contact.
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
The functional derivative $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
in the above formula must be computed with the asymmetric formulation
for the fundamental measures.  In practice, we use the approximate
White Bear functional to find an approximate expression for
$g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
expression will involve convolutions of derivatives of the free
energy, making this formulation computationally more expensive than
the free energy itself.

In addition to the different densities in the denominators of Equations 
\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
between the treatment of the A and S cases in the derivatives 
$\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
that we use involve a number of weighted densities and for the derivatives 
of these we use two different conceptual approaches.  In the assymetric case, 
a change in $\sigma(\mathbf{r})/2$ will refer to a change in the radius of a sphere 
that is located at the position $\mathbf{r}$.  The weighted densities that are 
affected by such a change will be those that are located a distance $\sigma/2$ away 
from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.

@@

These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the
derivation of the correlatation function results when we recognize that
instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$
as in the assymmetric case we will divide by $n_0(\mathbf{r})^2$
since we have symmetry around the point $\mathbf{r}$.

\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^S}\\
  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \\
  g_\sigma^S(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
\end{align}
where the functional derivative $\frac{\delta A_{HS}}{\delta
  \sigma(\mathbf{r})}$ in this case is computed with the symmetric
formulation of the fundamental measures.  Complete expressions for
$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\

@@

In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.  We break these terms up as follows:

@@

We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.  We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.  Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.  These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^A(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position of the sphere in
question---which is to say, the location where the density itself is
evaluated.  This is the most natural way to construct such a
functional, in which the radius of spheres is a function of position.
Thus we define $n_3^A$ as
\begin{align}
  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr'')/2) d\rr''
\end{align}
and we define the other weighted densities similarly.  In order to
compute derivatives of the free energy with respect to $R(\rr)$, we
need to find $g_\sigma^A$.  In the case of $n_3^A$, we need
\begin{align}
  \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
  \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
   &= n (\rr) \delta(|\rr' - \rr| - \sigma(\rr)/2)
\end{align}
Similarly, for $n_2^A$, we find
\begin{align}
  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) d \rr''\\
  \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
\end{align}
This is somewhat more challenging, since it involves an unfamiliar
function, the derivative the Dirac delta function $\delta'$.
%
The derivatives of the remaining scalar densities can be reduced to
sums of the terms above:
\begin{align}
  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
  \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
  &= -\frac{n(\rr)}{4\pi
    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
  -
  \frac{n(\rr)}{2\pi
    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2)
\end{align}
and
\begin{align}
  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
  \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
  &= -\frac{n(\rr)}{2\pi
    \sigma(\rr)^2}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
  -
  2 \frac{n(\rr)}{\pi
    \sigma(\rr)^3}\delta(|\rr'-\rr| - \sigma(\rr)/2)
\end{align}
The vector-weighted densities have no additional occurrences of $R$,
their functional derivatives come out similarly:
\begin{align}
  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(|\rr' - \rr''| - \sigma(\rr'')/2)
    \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
    \frac{\rr-\rr'}{|\rr-\rr'|}\\
\end{align}
and analogously for $\mathbf{n}_{V1}^A$:
\begin{align}
  \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
  \delta(|\mathbf{r'}-\rr''| - \sigma(\rr'')/2) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
  &= \frac{n(\rr)}{4\pi
    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
  +
  \frac{n(\rr)}{2\pi
    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
\end{align}
\fixme{double-check the directions of all the vector quantities...}

@@

The derivatation of the asymmetrically averaged correlation function
$g_\sigma^S(\rr)$ requires that we express the free energy in terms of
a set of weighted densities in which the radius of the weighting
kernel is determined according to the position at which the weighted
density is being evaluated.  This is less intuitive than the $A$ case,
since \emph{nothing is there}!  However, in practice it corresponds to
the picture shown in Figure~\ref{fig:contact}, which involves
extending the radius for spheres just at the position touching a
given spot, where we are evaluating the correlation function.  We
define the weighted density $n_3^S$ as:
\begin{align}
  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr')/2) d\rr''
\end{align}
In the symmetric case, the functional derivative is strikingly simple:
\begin{align}
   \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
   \delta(\rr'-\rr) \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
   \delta(\rr'-\rr) d\rr''
   \\
   &= \frac 12 \delta(\rr'-\rr) n_2^S(\rr')
\end{align}
The functional derivative does \emph{not} eliminate the integral,
which simply changes from one weighted density to another.  The
corresponding term in the free energy functional, once integrated,
takes a simple form:
\begin{align}
  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
  \frac{\delta n_3^{S}(\rr')}{\delta \sigma(\rr)} d\rr'
  &= \frac 12 \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
  n_2^S(\rr') d\rr' \\
  &= \frac 12 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
\end{align}
which means that we can compute $g_\sigma^S(\rr)$ using only
convolutions of the density, in strong contrast with
$g_\sigma^A(\rr)$, which requires convolutions of derivatives of the
free energy density.

The functional derivative of $n_2^S(\rr')$ is quite similar:
\begin{align}
  n_2^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf
  r''\\
  \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \frac 12 \int n(\rr'')
  \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
\end{align}
where we recognize that the convolution---while not one of the standard
FMT weighted density---is a simple convolution of the density.  We can
go on to compute the derivatives of the remaining weighted densities:
\begin{align}
  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
    \frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi \sigma(\rr)}
    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
    + \frac 12 n_0^s(\rr') \right)
\end{align}
\begin{align}
  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
    \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
    &= -\delta(\rr-\rr') \left( \frac 12 \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
    +
    \frac{2}{\sigma} n_0^S(\rr') \right)
\end{align}
\begin{align}
  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
  \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
  \frac 12 \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
  \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
\end{align}
\begin{align}
 \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''
   \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
 \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
 &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi \sigma(\rr)}
   \delta'(|\rr'-\rr''| - \sigma(\rr')/2) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
   + \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
\end{align}
